---
title: "FML Assignment 1"
output: html_document
date: "2024-01-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing the Datasets

```{r}
ds_salaries <- read.table("C:/Users/jackd/OneDrive/Desktop/R Projects/Fundamentals of Machine Learning/Datasets/ds_salaries.csv", header = TRUE, sep = ",")
```

## Descriptive Statistics

Average salary for all data science job titles:

```{r}
mean(ds_salaries$salary_in_usd)
```

Percentage of job titles.

```{r}
job_title_count <- table(ds_salaries$job_title) # calculate frequency of job titles
job_title_percentage <- prop.table(job_title_count) * 100 # calculate the proportion each job title makes of the whole and multiplies by 100 to find the percent
df_job_title_percentage <- data.frame(percentage = job_title_percentage) # converts back to a data.frame 
df_job_title_percentage <- df_job_title_percentage[order(-df_job_title_percentage$percentage.Freq), ] # orders the data.frame from large to small
print(df_job_title_percentage)
```

## Transform

Average salary for each job title:

```{r}
average_salary <- aggregate(ds_salaries$salary_in_usd ~ ds_salaries$job_title, FUN = mean, data = ds_salaries) # transforms and finds the mean of salaries grouped by job titles

print(average_salary)
```

Highest paid job title on average

```{r}
average_salary[which.max(average_salary$`ds_salaries$salary_in_usd`),] # finds the row where the maximum value resides and then selects said row
```

Looks like I need to become a Data Analytics Lead

## Plot

Lets make a Pie chart to see the frequency of each experience level

```{r}
experience_freq <- table(ds_salaries$experience_level) # calculates frequency of each experience level 
pie(experience_freq, main = "Frequency of Experience Levels") 
```

The only two variables which I can create a scatter plot from are the salary_in_usd and salary variable. So, even though it would not make sense to visualize these I will demonstrate how a scatter plot could be created to analyze the "Data Analyst" job title.

```{r}
options(scipen = 999) # disables scientific notation from the axis scales for readability 
data_analysts <- ds_salaries[ds_salaries$job_title == "Data Analyst", ] # filters ds_salaries by the job title "Data Analyst"
plot(data_analysts$salary, data_analysts$salary_in_usd) # plots salary by salary in usd
```
