---
title: "FML Assignment 1"
output: html_document
date: "2024-01-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing the Datasets

```{r}
ds_salaries <- read.table("C:/Users/jackd/OneDrive/Desktop/R Projects/Fundamentals of Machine Learning/Datasets/ds_salaries.csv", header = TRUE, sep = ",")
```

## Descriptive Statistics

Average salary for all data science job titles:

```{r}
mean(ds_salaries$salary_in_usd)
```

Number of unique job titles

```{r}
length(unique(ds_salaries$job_title))
```

## Transform

Average salary for each job title:

```{r}
average_salary <- aggregate(ds_salaries$salary_in_usd ~ ds_salaries$job_title, FUN = mean, data = ds_salaries)

print(average_salary)
```

Highest paid job title on average

```{r}
average_salary[which.max(average_salary$`ds_salaries$salary_in_usd`),]
```

Looks like I need to become a Data Analytics Lead

## Plot

Lets make a Pie chart to see the frequency of each experience level

```{r}
experience_freq <- table(ds_salaries$experience_level)
pie(experience_freq, main = "Frequency of Experience Levels")
```

The only two variables which I can create a scatterplot from are the salary_in_usd and salary variable. So even though it would not make sense to visualize these I will demonstrate a scatterplot.

```{r}
options(scipen = 999)
data_analysts <- ds_salaries[ds_salaries$job_title == "Data Analyst", ]
plot(data_analysts$salary, data_analysts$salary_in_usd)
```
